---
title: "Live Assignment 11"
author: "Chris Donlan"
date: "November 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
```

Load ffp and it's dependencies 
```{r }
  library(fpp) 
```

<<todo>> add information about data <<todos>>
Load in ukcars dataset which is inlcuded in fpp library
```{r }
  data(ukcars)
```

Print out str and first few row of ukcars
```{r}
  str(ukcars)
  head(ukcars)
```

## Part A

Plot ukcars time-series
```{r }
  plot(ukcars)
```

It's difficult from the plot to identiy any seasonal trends. There is clear a downward trend from the starting year, 1977, to around 1981, at which the trend changes direction. This trends continues tunil around the year 2000, at which it changes direction towards a downward trend. There are clearly season dips in the series, but the pravaliing upward tred makes it difficult to identify.

## Part B
```{r }
  fitcars <- decompose(ukcars)
  plot(fitcars)
```

Using a classical decomposition of additive method we are able to larger trend-cycle and seasonal indicies more easily. For trends, we see a plot that is similar to the plot from part A, but smoothed out. We see a turn down at the start of the data, 1977, a change up in 1980, another slight change down in the late 90s, which levels off in the early 2000s.

For season indicies, the seasonal plot indicates a yearly cycle pattern. The seasonal cycle fluxuates between 20 and -50.

## Part C
Yes, there are clearly small seasonal turns in direction identifiable in plot from part A. Although, it's difficult to identify cycle and indicies. 

## Part D
Compute and plot the seasonally adjusted data. 
```{r }
fitcarsseason <- seasadj(fitcars)
plot(fitcarsseason)
```

## Part E
```{r}
  #Get length of vector
  length(fitcarsseason)
  
  #Change value near the middle of series to 500
  originalval <- ukcars[60]
  ukcars[60] <- 500
  
  #Recompute seasonal adjustment and plot
  fitcarsseason <- seasadj(decompose(ukcars))
  plot(fitcarsseason)
  
  #change it back to the original value
  ukcars[60] <- originalval
  fitcarsseason <- seasadj(decompose(ukcars))
```
The outliers change the plot dramatically. Even with smoothing, decompositon, and seasonal adjustment, the exsistent of the outliers cauases a significant change in direction of the trend.


## Part F
Update value in the begginging 
```{r}
  ukcars.partf.begin <- ukcars
  ukcars.partf.begin[5] <- 500
  plot(seasadj(decompose(ukcars.partf.begin)))
```
Update value towards the end of the series
```{r}
  ukcars.partf.end <- ukcars
  ukcars.partf.end[95] <- 500
  plot(seasadj(decompose(ukcars.partf.end)))
```

It does effect the the seasonly adjust plot of the outlier appears at the begining or end vs the middle.  

## Part G



